# **Exploration**

## Read data

The code chunk below reads in the final project data.  

```{r, read_final_data}
df <- readr::read_csv("fall2022_finalproject.csv", col_names = TRUE)
```


```{r, show_data_glimpse}
df %>% glimpse()
```

## Modify data

### Transform output
```{r, trans_logit}
df <- df %>% mutate(y = boot::logit(output), outcome = ifelse(output < 0.33, 'event', 'non_event'), outcome = factor(outcome, levels = c("event", "non_event"))) %>%  glimpse()
```

### Add derived features
```{r, add_derived}
df_derived <- df %>% mutate(x5 = 1-x1-x2-x3-x4, w = x2 / (x3 + x4), z = (x1 + x2) / (x4 + x5), t = v1 * v2) %>%  glimpse()
```

## Visualize the data

### Visualize features
```{r, vis_distribution}
h_x1 <- df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = x1), bins=20)
h_x2 <- df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = x2), bins=20)
h_x3 <- df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = x3), bins=20)
h_x4 <- df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = x4), bins=20)
h_x5 <- df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = x5), bins=20)
h_v1 <- df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = v1), bins=20)
h_v2 <- df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = v2), bins=20)
h_v3 <- df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = v3), bins=20)
h_v4 <- df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = v4), bins=20)
h_v5 <- df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = v5), bins=20)
h_w <- df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = w), bins=20)
h_z <- df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = z), bins=20)
h_t <- df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = t), bins=20)
ggarrange(h_x1, h_x2, h_x3, h_x4, h_x5, h_v1, h_v2, h_v3, h_v4, h_v5, h_w, h_z, h_t, nrow = 4, ncol = 4) +
  theme_bw()
```
```{r, vis_distribution_m}
df_derived %>% 
  ggplot() +
  geom_bar(aes(x = m)) +
  theme_bw()
```

### Visualize outputs
```{r, vis_distribution_output}
h_output <- df %>% 
  ggplot() +
  geom_histogram(aes(x = output), bins = 20)
h_y <- df %>% 
  ggplot() +
  geom_histogram(aes(x = y), bins = 20)
ggarrange(h_output, h_y, ncol = 2) +
  theme_bw()
```

## Distribution with respect to `m`

### Features with respect to `m`
```{r, m_dis_x1}
df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = x1), bins=20) +
  facet_wrap(~m) +
  theme_bw()
```
```{r, m_dis_x2}
df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = x2), bins=20) +
  facet_wrap(~m) +
  theme_bw()
```
```{r, m_dis_x3}
df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = x3), bins=20) +
  facet_wrap(~m) +
  theme_bw()
```
```{r, m_dis_x4}
df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = x4), bins=20) +
  facet_wrap(~m) +
  theme_bw()
```
```{r, m_dis_x5}
df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = x5), bins=20) +
  facet_wrap(~m) +
  theme_bw()
```
```{r, m_dis_v1}
df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = v1), bins=20) +
  facet_wrap(~m) +
  theme_bw()
```
```{r, m_dis_v2}
df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = v2), bins=20) +
  facet_wrap(~m) +
  theme_bw()
```
```{r, m_dis_v3}
df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = v3), bins=20) +
  facet_wrap(~m) +
  theme_bw()
```
```{r, m_dis_v4}
df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = v4), bins=20) +
  facet_wrap(~m) +
  theme_bw()
```
```{r, m_dis_v5}
df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = v5), bins=20) +
  facet_wrap(~m) +
  theme_bw()
```
```{r, m_dis_w}
df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = w), bins=20) +
  facet_wrap(~m) +
  theme_bw()
```
```{r, m_dis_z}
df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = z), bins=20) +
  facet_wrap(~m) +
  theme_bw()
```
```{r, m_dis_t}
df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = t), bins=20) +
  facet_wrap(~m) +
  theme_bw()
```

<font color=Blue>
There are no obvious differences in input values based on the discrete groups.
</font> 

### Output with respect to `m`
```{r, m_dis_output}
df_derived %>% 
  ggplot() +
  geom_histogram(aes(x = output), bins=20) +
  facet_wrap(~m) +
  theme_bw()
```

<font color=Blue>
There are no obvious differences in `output` based on the discrete groups.
</font> 

## Relation between features
```{r, cor_feature}
df_derived %>% 
  select(starts_with('x'), starts_with('v'), c(w, z, t)) %>% 
  cor() %>% 
  corrplot::corrplot(type = 'upper')
```

<font color=Blue>
The derived features of course are correlated to the inputs. For the inputs part, `x1,x2,x3` have some correlations.
</font> 

## Relation between output and features
```{r, relation}
p_x1 <- df_derived %>% ggplot(mapping = aes(x = x1, y = y)) +
  geom_point(alpha = 0.33)
p_x2 <- df_derived %>% ggplot(mapping = aes(x = x2, y = y)) +
  geom_point(alpha = 0.33)
p_x3 <- df_derived %>% ggplot(mapping = aes(x = x3, y = y)) +
  geom_point(alpha = 0.33)
p_x4 <- df_derived %>% ggplot(mapping = aes(x = x4, y = y)) +
  geom_point(alpha = 0.33)
p_x5 <- df_derived %>% ggplot(mapping = aes(x = x5, y = y)) +
  geom_point(alpha = 0.33)
p_v1 <- df_derived %>% ggplot(mapping = aes(x = v1, y = y)) +
  geom_point(alpha = 0.33)
p_v2 <- df_derived %>% ggplot(mapping = aes(x = v2, y = y)) +
  geom_point(alpha = 0.33)
p_v3 <- df_derived %>% ggplot(mapping = aes(x = v3, y = y)) +
  geom_point(alpha = 0.33)
p_v4 <- df_derived %>% ggplot(mapping = aes(x = v4, y = y)) +
  geom_point(alpha = 0.33)
p_v5 <- df_derived %>% ggplot(mapping = aes(x = v5, y = y)) +
  geom_point(alpha = 0.33)
p_w <- df_derived %>% ggplot(mapping = aes(x = w, y = y)) +
  geom_point(alpha = 0.33)
p_z <- df_derived %>% ggplot(mapping = aes(x = z, y = y)) +
  geom_point(alpha = 0.33)
p_t <- df_derived %>% ggplot(mapping = aes(x = t, y = y)) +
  geom_point(alpha = 0.33)
p_m <- df_derived %>% ggplot(mapping = aes(x = m, y = y)) +
  geom_point(alpha = 0.33)
ggarrange(p_x1, p_x2, p_x3, p_x4, p_x5, nrow = 3, ncol=2) +
  theme_bw()
ggarrange(p_v1, p_v2, p_v3, p_v4, p_v5, nrow = 3, ncol=2) +
  theme_bw()
ggarrange(p_w, p_z, p_t, p_m, nrow = 2, ncol=2) +
  theme_bw()
```

## Relation between binary outcome and features
```{r, binary_graph}
b_x1 <- df_derived %>% 
  ggplot(mapping = aes(x = x1, y = outcome)) +
  geom_jitter(height = 0.02, width = 0, alpha=0.15) +
  theme_bw()
b_x2 <- df_derived %>% 
  ggplot(mapping = aes(x = x2, y = outcome)) +
  geom_jitter(height = 0.02, width = 0, alpha=0.15) +
  theme_bw()
b_x3 <- df_derived %>% 
  ggplot(mapping = aes(x = x3, y = outcome)) +
  geom_jitter(height = 0.02, width = 0, alpha=0.15) +
  theme_bw()
b_x4 <- df_derived %>% 
  ggplot(mapping = aes(x = x4, y = outcome)) +
  geom_jitter(height = 0.02, width = 0, alpha=0.15) +
  theme_bw()
b_x5 <- df_derived %>% 
  ggplot(mapping = aes(x = x5, y = outcome)) +
  geom_jitter(height = 0.02, width = 0, alpha=0.15) +
  theme_bw()
b_v1 <- df_derived %>% 
  ggplot(mapping = aes(x = v1, y = outcome)) +
  geom_jitter(height = 0.02, width = 0, alpha=0.15) +
  theme_bw()
b_v2 <- df_derived %>% 
  ggplot(mapping = aes(x = v2, y = outcome)) +
  geom_jitter(height = 0.02, width = 0, alpha=0.15) +
  theme_bw()
b_v3 <- df_derived %>% 
  ggplot(mapping = aes(x = v3, y = outcome)) +
  geom_jitter(height = 0.02, width = 0, alpha=0.15) +
  theme_bw()
b_v4 <- df_derived %>% 
  ggplot(mapping = aes(x = v4, y = outcome)) +
  geom_jitter(height = 0.02, width = 0, alpha=0.15) +
  theme_bw()
b_v5 <- df_derived %>% 
  ggplot(mapping = aes(x = v5, y = outcome)) +
  geom_jitter(height = 0.02, width = 0, alpha=0.15) +
  theme_bw()
b_w <- df_derived %>% 
  ggplot(mapping = aes(x = w, y = outcome)) +
  geom_jitter(height = 0.02, width = 0, alpha=0.15) +
  theme_bw()
b_z <- df_derived %>% 
  ggplot(mapping = aes(x = z, y = outcome)) +
  geom_jitter(height = 0.02, width = 0, alpha=0.15) +
  theme_bw()
b_t <- df_derived %>% 
  ggplot(mapping = aes(x = t, y = outcome)) +
  geom_jitter(height = 0.02, width = 0, alpha=0.15) +
  theme_bw()
ggarrange(b_x1, b_x2, b_x3, b_x4, b_x5, nrow = 3, ncol = 2)
ggarrange(b_v1, b_v2, b_v3, b_v4, b_v5, nrow = 3, ncol = 2)
ggarrange(b_w, b_z, b_t, nrow = 2, ncol = 2)
df_derived %>% 
  ggplot(mapping = aes(x = m)) +
  geom_bar(mapping = aes(fill = outcome),
           position = 'fill') +
  labs(y = 'group proportion') +
  theme_bw()
```

